### 2019-10-17 thesis_template/3_Body/body.tex
changed latex stracture related work is added

\chapter{Body of my thesis}
\label{chapter:body}
\thispagestyle{myheadings}

% set this to the location of the figures for this chapter. it may
% also want to be ../Figures/3_Body/ or something. make sure that
% it has a trailing directory separator (i.e., '/')!
\graphicspath{{3_Body/Figures/}}

\section{Some results}
\label{sec:results}

Here goes all the important stuff, likely with a lot of graphics like this:

\begin{figure}[htb]
  \begin{minipage}[t]{0.49\linewidth}\centering
    \includegraphics[width=7cm]{figure_sampling_view1.eps}
    \medskip
    \centerline{(a)}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.49\linewidth}\centering
    \includegraphics[width=7cm]{figure_sampling_view2.eps}
    \medskip
    \centerline{(b)}
  \end{minipage}
  \caption{Assignment of single-view intensities to RGB components: (a) view
    \#1; and (b) view \#2. }
  \label{fig:Sampling}
\end{figure}

You will also be using a lot of citations. Here is the format required in the dissertation: \cite{lamport1985:latex},\cite{Debr01}.

In all likelihood, you will need to insert tables. See one example on the next page.
\clearpage

\begin{table}[h]
	\caption{Absolute disparity error per pixel for the test data from
		Fig.~\ref{fig:Sampling} and different parameter values. In each experiment one
		parameter is adjusted while other parameters are unchanged.} 
	\centering
	\begin{minipage}[b]{0.30\linewidth}
		\centerline{$\eta=6000$, $\mu=2000$}\smallskip
		\centering
		\begin{tabular}{ccc}
			\hline
			$K$ & $u_1$ & $u_2$\\
			\hline
			3   & 0.52 &0.46\\
			7   & 0.47 &0.43\\
			10  & 0.35 &0.36\\
			12  & 0.37 &0.36\\
			\hline
		\end{tabular}
	\end{minipage}
	%
	\begin{minipage}[b]{0.34\linewidth}
		\centerline{$K=10$, $\mu=2000$}\smallskip
		\centering
		\begin{tabular}{ccc}
			\hline
			$\eta$ & $u_1$ & $u_2$\\
			\hline
			1000&0.54& 0.45\\
			3000&0.43& 0.40\\
			6000&0.35& 0.36\\
			9000&0.37& 0.37\\
			\hline
		\end{tabular}
	\end{minipage}
	%
	\begin{minipage}[b]{0.32\linewidth}
		\centerline{$K=10$, $\eta=6000$}\smallskip
		\centering
		\begin{tabular}{ccc}
			\hline
			$\mu$ & $u_1$ & $u_2$\\
			\hline
			100 &1.00&1.16\\
			1000&0.53&0.47\\
			2000&0.35&0.36\\
			3000&0.44&0.43\\
			\hline
		\end{tabular}
	\end{minipage}
	%
	\label{tab:Parameters}
\end{table}

Of course, there must be a Table of Contents at the beginning of the thesis.