### 2020-03-10 thesis_template/2_RelatedWork/model/tree_aggregate_experiment.tex
added GC tuning section

Tree-aggregate is a communication patten heavily used for Machine Learning algorithm in Spark (MLlib). 
In the traditional aggregation function in Spark, results of aggregation in all executor clusters are sent to the driver. 
That is why this operation suffers from the CUP cost in merging partial results and the network bandwidth limit.
Tree-aggregate is a communication pattern which overcomes these problems by breaking aggregate operation in multi-level represented like tree structure.

In multi-thread situation, tree-aggregate is just fork-join thread algorithm. We already implemented Merge-Sort algorithm with fork-join. 
If we implement tree-aggregate with multi-thread, the difference is only aggregation and sorting.