### 2020-04-03 thesis_template/2_RelatedWork/relatedWork.tex

\chapter{Related Work}
\label{chapter:relatedwork}
\thispagestyle{myheadings}

\graphicspath{{2_RelatedWork/Figures/}}

\label{sec:history}
\input{2_RelatedWork/model/spark}
\clearpage

\section{Serialization}
\label{sec:history}
\input{2_RelatedWork/model/serialization}
\clearpage

\section{LLVM}
\label{sec:history}
\input{2_RelatedWork/model/llvm}

\section{Todos}
\label{sec:history}
\begin{itemize}
    \item Document for Serialization
    \item Document for Region Based Memory Management
    \item Document for LLVM in more detail.
    \item Document for Rust in more detail.
\end{itemize}