### 2020-04-17 thesis_template/3_Body/body.tex
documented concept

\chapter{Experiment}
\label{chapter:body}
\thispagestyle{myheadings}

% set this to the location of the figures for this chapter. it may
% also want to be ../Figures/3_Body/ or something. make sure that
% it has a trailing directory separator (i.e., '/')!
\graphicspath{{3_Body/Figures/}}

\section{Concept}
\label{sec:history}
\input{3_Body/chapter/concept}
\clearpage

\section{Evaluation}
\label{sec:history}
\input{3_Body/chapter/evaluation}
\clearpage