### 2020-04-17 thesis_template/3_Body/chapter/evaluation.tex

\subsection{Experiment for Accessing Object with Different Variable Type}
\label{sec:history}
\input{3_Body/model/access_own_reference_slice}
\clearpage

\subsection{Assessment of different reference methods in Rust}
\label{sec:history}
\input{3_Body/model/drop_ref_rc}
\clearpage

\subsection{Experiment for Merge-sort}
\label{sec:history}
\input{3_Body/model/merge_sort_experiment}
\clearpage

\subsection{Experiment for Tree-aggregation}
\label{sec:history}
\input{3_Body/model/tree_aggregate_experiment}
\clearpage

\subsection{Elements Copy and Insertion into Size-initialized Vector in Rust.}
\label{sec:history}
\input{3_Body/model/element_copy_initialization}