### 2020-04-22 thesis_template/2_RelatedWork/relatedWork.tex

\chapter{Related Work}
\label{chapter:relatedwork}
\thispagestyle{myheadings}

\graphicspath{{2_RelatedWork/Figures/}}

\section{Hadoop MapReduce to Spark}
\label{sec:history}
\input{2_RelatedWork/model/hadoop_to_spark}

\section{Apache Spark}
\label{sec:history}
\input{2_RelatedWork/model/spark}

\section{Application to System Language}
\label{sec:history}
\input{2_RelatedWork/model/application_to_system_language}

\section{Rust Memory Management}
\label{sec:history}
\input{2_RelatedWork/model/rust_memory_management}

\section{LLVM}
\label{sec:history}
\input{2_RelatedWork/model/llvm}

\section{Serialization}
\label{sec:history}
\input{2_RelatedWork/model/serialization}

\section{Todos}
\label{sec:history}
\begin{itemize}
    \item Ball-tree and KD-tree
    \item Rust mutability and lifetime
    \item Document for Serialization
    \item Document for Region Based Memory Management
    \item Document for LLVM in more detail.
    \item Document Reference Counting and Atomic Reference Counting
\end{itemize}
\clearpage