### 2020-04-23 thesis_template/Appendix/Appendix.tex

\graphicspath{{Appendix/Figures/}}

\chapter{Operating Systems}
\label{appendix}
\thispagestyle{myheadings}
\input{Appendix/model/os}

\chapter{Linear Algebra Computation}
\label{appendix}
\thispagestyle{myheadings}
\input{Appendix/model/linalg}