### 2020-04-26 thesis_template/0_Prelim/abs.tex
labled documentation

% ABSTRACT

Many existing Big Data processing tools are developed with application languages. 
Dataflow platform such as Hadoop MapReduce,  Apache Spark, and Apache Flink 
developed on Java Virtual Machine (JVM). JVM abstracts hardware and memory management from the developer. 
This eases development processes and distribution of these tools by eliminating manual memory management and 
platform dependency.

However, the automated memory strategy, such as Garbage Collection (GC), may lead significant overhead in Big Data processing. 
Apache Spark and Apache Flink use complex objects to manipulate and transfer tremendous amount of data. 
Generating many of these complex objects in memory forces GC to rearrange the objects in memory frequently waisting computation. 
Keeping these objects in memory for long period of time may trigger stop-the-world where the JVM stops application from running to
execute a GC.

Considered these problems in memory management in JVM, developing Big Data processing tools with system language can be the solution.
By using system language, a developer has control on memory management, so that one can implement systems with better optimized memory management strategies.
We selects Rust as good candidate for development of Big Data processing tools, due to its ability to write memory-safe and fearless concurrent codes.

To implement optimal memory management in Rust for such tools, there may be variety of strategies and tools. 
In this dissertation, we conduct experiments to examine the best memory management strategies in Rust for Big Data processing.