### 2020-04-26 thesis_template/2_RelatedWork/relatedWork.tex
document updated

\chapter{Related Work}
\label{chapter:relatedwork}
\thispagestyle{myheadings}

\graphicspath{{2_RelatedWork/Figures/}}

\section{Operating Systems}
\label{sec:os}
\input{2_RelatedWork/model/os}

\section{Linear Algebra Computation}
\label{sec:linalg}
\input{2_RelatedWork/model/linalg}

\section{Hadoop MapReduce to Spark}
\label{sec:hadoopspark}
\input{2_RelatedWork/model/hadoop_to_spark}

\section{Apache Spark}
\label{sec:spark}
\input{2_RelatedWork/model/spark}

\section{Application to System Language}
\label{sec:appsys}
\input{2_RelatedWork/model/application_to_system_language}

\section{Rust Memory Management}
\label{sec:rustmem}
\input{2_RelatedWork/model/rust_memory_management}

\section{LLVM}
\label{sec:llvm}
\input{2_RelatedWork/model/llvm}

% \section{Serialization}
% \label{sec:history}
% \input{2_RelatedWork/model/serialization}

% \section{Todos}
% \label{sec:history}
% \input{2_RelatedWork/model/todo}

\clearpage