### 2020-04-26 thesis_template/3_Body/body.tex

\chapter{Conceptual Design of Experience}
\label{chapter:conept}
\thispagestyle{myheadings}

% set this to the location of the figures for this chapter. it may
% also want to be ../Figures/3_Body/ or something. make sure that
% it has a trailing directory separator (i.e., '/')!
\graphicspath{{3_Body/Figures/}}
\input{3_Body/chapter/concept}
\clearpage

\chapter{Evaluation Result}
\label{chapter:evaluation}
\thispagestyle{myheadings}
\input{3_Body/chapter/evaluation}
\clearpage