### 2020-04-26 thesis_template/3_Body/chapter/evaluation.tex

\section{Experimental Set and Detail}
\label{sec:eval_setdetail}
\input{3_Body/model/experimental_set_detail}
\clearpage

\section{Experiment 1: Accessing Object with Different Variable Type}
\label{sec:eval_diffval}
\input{3_Body/model/access_own_reference_slice}
\clearpage

\section{Experiment 2: Assessment of different reference methods in Rust}
\label{sec:eval_refcount}
\input{3_Body/model/drop_ref_rc}
\clearpage

\section{Experiment 2: Merge-sort}
\label{sec:eval_sort}
\input{3_Body/model/merge_sort_experiment}
\clearpage

\section{Experiment 4: Tree-aggregation}
\label{sec:eval_treeagg}
\input{3_Body/model/tree_aggregate_experiment}
\clearpage

% \section{Elements Copy and Insertion into Size-initialized Vector in Rust.}
% \label{sec:history}
% \input{3_Body/model/element_copy_initialization}