### 2020-04-26 thesis_template/3_Body/model/experimental_set_detail.tex

\subsection{Wikipedia Data Sets}
 \label{sec:concept_dataset}
 Wikipedia page data sets are used to perform document classification with KNN. 
 \(10^5\) pages are used for training data set, and 18724 pages are used for target.

 \subsection{Experimental Details}
 \label{sec:concept_expdetail}
 Our experiments are run on three types of VM instances on Google Cloud Platform: 
 n1-standard-1 which has 1 vCPU, 3.75 GB RAM, and 10 GB Standard persistent disk, 
 n1-standard-4 which has 4 vCPU, 15 GB RAM, and 10 GB Standard persistent disk, 
 n1-standard-8 which has 8 vCPU, 30 GB RAM, and 10 GB Standard persistent disk,  
 All of the experiments described are performed 5 times. 
 In this thesis, we present the average of 5 separate runs for each experiment.