### 2020-04-26 thesis_template/Appendix/Appendix.tex

\graphicspath{{Appendix/Figures/}}

% \chapter{Operating Systems}
% \label{appendix}
% \thispagestyle{myheadings}
% \input{Appendix/model/os}

\chapter{Linear Algebra Computation}
\label{appendix}
\thispagestyle{myheadings}
\input{Appendix/model/linalg}