### 2020-04-27 thesis_template/2_RelatedWork/model/intro.tex
related intro

This chapter describes background knowledge that we have studied for our research.
In Section~\ref{sec:os}, some of topics in Operating Systems are described focusing on memory management in operating systems.
Knowledge of operating systems helps us to understand how operating systems handle memory management and 
how other technologies takes advantage of them. Even though this dissertation does not focus technologies used for Linear Algebra Computation, 
there are some active researches in technologies in Linear Algebra Computation for Big Data processing. 
It worths to study these concepts explained in Section~\ref{sec:linalg}. 
In Section~\ref{sec:hadoopspark}, major Big Data processing tools and improvement of these tools are introduced. 
Then, Section~\ref{sec:spark} details technologies used for Apache Spark, which is one of the most popular tools for Big Data processing.
In Section~\ref{sec:appsys} and Section~\ref{sec:rustmem}, the advantages of using System Language and Rust for such tools are discussed respectively.
Finally, Section~\ref{sec:llvm} describes how LLVM thrives computer language development.