### 2020-04-27 thesis_template/2_RelatedWork/model/summary.tex
added intro and summary for each chapter

In this chapter, we cover some of main concepts to understand our research. 
The topics of Operating Systems and Linear Algebra Computation are not directly relevant to our research, 
but it is worth to cover these topics for deep analysis of experimental result. 
The advantage and disadvantage of modern Big Data processing tools are discussed and 
some of solutions are introduced. Finally, how Rust can be a good candidate for development of Big Data processing tools and 
details of its technologies are described. In next chapter, we discuss main concepts needed to understand theoretical assumption of our experiments.