### 2020-05-06 thesis_template/2_RelatedWork/model/gctuning.tex
need object size and conclusion documentation

There are many different ways that one can improve the performance of GC in Java. One of these is for example avoiding pointer-based data structures, such as HashMap and LinkedList. 
These objects have a "wrapper" object for each entry so that number of object tends to be larger than when when an array is used.

Caching serialized object in memory also reduce number of object and memory usage since the set of objects become a byte or binary array. 
Spark SQL applications use DataFrames\cite{DBLP:conf/sigmod/ArmbrustXLHLBMK15}, 
whose intermediate data are managed by an optimized memory manager named Tungsten.
Tungsten stores the intermediate data in a serialized binary form and performs aggregation functions directly on he serialized objects.
Therefore, the number of Java objects in memory is reduced and it reduces the DC frequency and object marking/sweeping.

In addition, developers can allocate data off-heap of JVM to avoid tracking by GC. 
Facade\cite{DBLP:conf/asplos/NguyenWBFHX15} proposed a compiler and runtime system to bound the number of in-memory data objects, through storing data in an off-heap region and 
manipulating the data with control interfaces.

Although these memory management solution for GC help developer improve performance of Spark applications, 
the effort to discover the best GC tuning afflicts developers.