### 2020-06-28 thesis_template/0_Prelim/abs.tex
corrected

% ABSTRACT

% Motivation
% - Faster computation and  saving computer resource is important
% - introduce Big Data tool can be improve
% Problem description
% - introduce Spark, MapReduce is implemented in Application Language.
% - some problems
% Concept
% - use of rust
% - what is core concept of experiment
% Evaluation
% - how to evaluate that
% - some notable result

Planning optimized memory management is critical for Big Data analysis tools to
perform faster runtime and efficient use of computation resources.
Modern Big Data analysis tools use application languages that abstract their
memory management so that developers do not have to pay extreme attention
to memory management strategies.

Many existing modern cloud-based data processing systems such as Hadoop, Spark or Flink use
Java Virtual Machine (JVM) and take full advantage of features such as automated memory management in JVM
including Garbage Collection (GC) which may lead to a significant overhead.
Dataflow-based systems like Spark allow programmers to define complex objects in a
host language like Java to manipulate and transfer tremendous amount of data.

% Generating many of these complex objects in memory forces GC to rearrange the objects in memory frequently waisting computation.
% Considered performance overhead in memory management in JVM makes system languages to be a better choice to
% develop systems for Big Data processing.

System languages like C++ or Rust seem to be a better choice to
develop systems for Big Data processing because they do not relay on JVM.
By using a system language, a developer has full control on the memory management.
% Therefore, one can implement systems with more optimized memory management strategies.
We found Rust programming language to be a good candidate due to its ability to write memory-safe and
fearless concurrent codes with its concept of memory ownership and borrowing.
Rust programming language includes many possible strategies to optimize memory management for Big Data processing including
a selection of different variable types, use of Reference Counting, and multithreading with Atomic Reference Counting.

In this thesis, we conducted an experimental study to assess how much these different memory management
strategies differ regarding overall runtime performance.
Our experiments focus on complex object manipulation and common Big Data processing patterns
with various memory management. Our experimental results indicate a significant
difference among these different memory strategies regarding data processing performance.