### 2020-06-28 thesis_template/0_Prelim/ack.tex

Foremost, I would like to show my appreciation to my advisor Prof. Kia Teymourian for his guidance, encouragement and support.
His expertise and insights were essential for this research work. His passion for research and education motivates me and his experience is instrumental to the completion of the thesis.

I am especially thankful to Prof. Eugene Pinsky for accepting reviewing of this thesis. 
I also thank Prof. Reza Rawassizadeh for review and advice on this thesis.

Finally, I would like to express special thanks to my parents for supporting this excellent educational opportunity.

\vskip 1in

\noindent
Janusz Konrad\\
Professor\\
ECE Department