### 2020-06-28 thesis_template/2_RelatedWork/model/summary.tex

In this chapter, we cover some of main concepts to understand our research. 
The topics of Operating Systems and Linear Algebra Computation are not directly relevant to our research, 
but it is worth covering these topics for deep analysis of experimental result. 
The advantage and disadvantage of modern Big Data processing tools are discussed and 
some solutions are introduced. Finally, how Rust can be a good candidate for development of Big Data processing tools and 
details of its technologies are described. In the next chapter, we discuss main concepts needed to understand theoretical assumptions of our experiments.