### 2020-06-28 thesis_template/3_Body/chapter/evaluation.tex

\section{Introduction}
\label{sec:eval_intro}
In this chapter, detailed experiment settings, executions and results are discussed.
Section~\ref{sec:eval_setdetail} details hardware configurations and datasets we use. 
There are 5 experiments conducted. In Section~\ref{sec:eval_diffval}, we discuss the experiment 
where we assess runtime to access object's fields with different variable types. 
Section~\ref{sec:eval_refcount} shows details to examine how behaviors of normal reference and Reference Count are different. 
Section~\ref{sec:eval_sort} describes our Merge-sort experiment to examine behavior of Atomic Reference Counting.
Two algorithms are used in Big Data processing and examined in our experiments: 
Tree-aggregate in Section~\ref{sec:concept_treeagg} and K-Nearest-Neighbors in Section~\ref{sec:concept_knn}.

\section{Experimental Set and Detail}
\label{sec:eval_setdetail}
\input{3_Body/model/experimental_set_detail}

\section{Experiment 1: Accessing Objects with Different Variable Types}
\label{sec:eval_diffval}
\input{3_Body/model/access_own_reference_slice}

\section{Experiment 2: Assessment of different reference methods in Rust}
\label{sec:eval_refcount}
\input{3_Body/model/drop_ref_rc}

\section{Experiment 3: Merge-sort}
\label{sec:eval_sort}
\input{3_Body/model/merge_sort_experiment}

\section{Experiment 4: Tree-aggregation}
\label{sec:eval_treeagg}
\input{3_Body/model/tree_aggregate_experiment}

\section{Experiment 5: K-Nearest-Neighbors}
\label{sec:eval_knn}
\input{3_Body/model/knn}

\section{Summary}
\label{sec:eval_summary}
In this chapter, we examine various memory management strategies for different algorithms in Rust programming.
There are many factors that might differ performance of algorithms: different variable type, Reference Counting, Atomic Reference Counting, and frequency of memory de/allocation.
We should select the best memory management strategies when developing Big Data processing tools. 
Some tips for the development are proven in our experiments. We conclude this thesis in Chapter~\ref{chapter:Conclusions}.

% \section{Elements Copy and Insertion into Size-initialized Vector in Rust.}
% \label{sec:history}
% \input{3_Body/model/element_copy_initialization}