### 2020-06-28 thesis_template/3_Body/model/experimental_set_detail.tex

We have used two different datasets. One is a real-world data set, Wikipedia page data set,
and another is a syntactic randomized generated data set based on Complex objects described in Section~\ref{sec:concept_compobj}. 
We use one machine with particular specifications.
\subsection{Wikipedia Data Sets}
 \label{sec:concept_dataset}
 Wikipedia page data sets are used to perform document classification with KNN. 
 We have separated the data set into training and test data set, 
 \(10^5\) pages are used for training data set, and 18724 pages are used for target.

 \subsection{Experimental Details}
 \label{sec:concept_expdetail}
 All experiments are run on VM instances on Google Cloud Platform, 
 n1-standard-8 which has 8 vCPU, 30 GB RAM, and 10 GB Standard persistent disk.
 In this thesis, we present the result of runtime as the average of 5 separate runs for each experiment.