### 2020-06-28 thesis_template/3_Body/model/merge_sort_experiment.tex

Sorting algorithm like merge-sort is a very ubiquitous algorithm in any computation. 
Especially in Big Data processing, merge-sort algorithms de/allocate contiguous memory for partitions of sorted vector multiple times.
We are interested in memory de/allocation patterns in merge-sort algorithm, because it generates many intermediate data structures and deletes them many times.

As explained in Experiment 2 in Section~\ref{sec:eval_refcount}, deletion of Rc has significant overhead than normal reference. 
By learning this result, our assumption here is that deletion of Arc also has overhead when we compare to normal reference. 
In many situations where a developer writes a multithreading code, Arc is used and the deletion happens multiple times. 
To assess runtime overhead of algorithms with Arc, we implement merge-sort algorithm in two different ways. 
One is sharing source vector with Arc. The other is passing reference of source vector to child thread.

Our merge-sort algorithms are implemented with recursion. For each call of recursive function, 
Arc or slice of the source vector needs to be passed and deleted when the function returns value. 
These merge-sort algorithms trigger large number of Arc or reference deletion proportional to the number of call recursive function.
Merge-sort algorithm can be separated in three phases: splitting phase, copying phase, and merging phase.

The splitting phase is merely acquiring index of range. At this phase, multiple threads are generated and Arc or reference of source vector are passed by calling recursive function. 
Copying phase occurs in the base case of the recursive call. The element in the source vector in a certain index is deep-copied into newly allocated vector.
At merge phase, merge function receives two sorted independent vectors and merges them into a single new vector.

We use scope method from crossbeam crate to perform multithread programming. 
Scoped thread can have reference to value from its parent thread by ensuring children threads are joined before their parent thread returns value. 
By using scoped thread, we can implement two merge-sort algorithms in an identical way except whether the function receives Arc or reference of source vector.
The representations of source vectors for each algorithm are shown in Figure~\ref{fig:source_merge_sort}.
The elements of source vector are CustomerOwned objects. We generate source vectors in sizes 4, 8, 12 and 16 million.
Finally, merge-sort is performed based on value of key field. The figure shows the result for runtime performance of merge-sort algorithms on difference sizes of vectors.

\begin{figure}[htb]
    \begin{lstlisting}
        // Source vector for algorithm with Arc.
        arr: Arc<VecDeque<T>>

        // Source vector for algorithm with reference (slice).
        arr: &[T]
    \end{lstlisting}
    \caption{Representation of Source Vector}
    \label{fig:source_merge_sort}
\end{figure}

\subsection{Result}
Figure~\ref{fig:source_merge_sort} shows the runtime performance of our merge-sort algorithm with specified Vec sizes. 
The blue and yellow bar charts represent the runtime performance of merge-sort algorithm using reference and Arc respectively. 
The result says that algorithms with Arc is about 21\% slower than algorithms with reference.
The memory usage for algorithms with 16 million Customer objects are about 26G bytes.

\begin{figure}[htb]
    \includegraphics[width=15cm]{rust_merge_sort.eps}
    \caption{Runtime of Sorting Elements of Customer Vector}
    \label{fig:merge_sort}
\end{figure}

\subsection{Discussion}
The reason why merge-sort algorithms with Arc is much slower than one with reference is 
the same for the reason why dropping Rc shows overhead compared to dropping reference.
Arc has to check the number of variable pointing to the actual content and decide 
deallocate the memory or not.

In addition, Arc uses atomic operations for reference counting. Atomic operations bring thread-safety, but they are more expensive than ordinary memory accesses.
Therefore, when sharing reference counting between threads is not required, using Rc is the recommended way \cite{RustArcPage}.

In situations like our experiment, normal reference can be used instead of Arc to share data between threads. 
This solution results in better runtime performance in our experiment. Therefore, one should use reference to share data among different threads 
whenever it is possible.