### 2020-10-11 thesis_template/0_Prelim/cv.tex
completed correction

\addcontentsline{toc}{chapter}{Curriculum Vitae}
\thispagestyle{empty}

\begin{center}
{\LARGE {\bf CURRICULUM VITAE}}\\
\vspace{0.5in}
\end{center}

\setcounter{section}{0}
\renewcommand*{\theHsection}{chX.\the\value{section}}

\section*{Contact}
\begin{itemize}
    \item Phone: +81-90-4148-4639 \textbullet{} City: Tokyo \textbullet{} Email: \href{golisaku4639@gmail.com}{golisaku4639@gmail.com} 
    \item GitHub: \href{https://github.com/ShinsakuOkazaki}{\textbf{github}.com/ShinsakuOkazaki} 
    \item LinkedIn\href{https://www.linkedin.com/in/shinsaku-okazaki-7b786b13b}{ \textbf{linkedin}.com/in/shinsaku-okazaki-7b786b13b}
\end{itemize}

\section*{Education}
\subsection*{Boston University, Boston, MA}
\begin{itemize}
    \item Program: Master of Science in Computer Information Systems (September 2018 - May 2020)
    \item Course Work: Software Design Pattern, Database, Big Data Analysis, Data Analysis
    \item Master Thesis: An Experimental Study of Memory Management with Rust Programming for Big Data Processing
\end{itemize}

\subsection*{Seikei University, Tokyo}
\begin{itemize}
    \item Program: Bachelor of Science in System Design (April 2013 - March 2018)
    \item Course Work: Image Processing, Control Engineering, Digital Signal Processing, Engineering Economics
\end{itemize}

\section*{Experience} 
\subsection*{Research Assistant, Boston University, September 2019 - May 2020}
\begin{itemize}
    \item  Improved run-time performance of Large Scale Sort Algorithm 30\% by implementing algorithm with careful Memory Management strategies in Rust.
    \item  Developed Merge-sort Algorithm and Tree-aggregate Algorithm with Multithreading using Rust on Cloud Instances using GCP.
    \item Implemented \textbf{Machine Learning} Algorithm with different memory management strategy to assess best workflow on \textbf{Natural Language Processing} with Rust.
\end{itemize}

\subsection*{Teaching Assistant, Boston University, September 2019 - May 2020}
\begin{itemize}
    \item  Developed a R package where student can take R programming exercises to improve student's learning experience.
    \item  Designed data collection feature by adding REST APIs using Flask and MongoDB.
    \item  Facilitated customization of software by reducing building process 60\% with automation using Bash script.
    \item  Collaborated with a team of 4 members to design flexible and robust software architecture.
\end{itemize}

\subsection*{Software Engineer Intern, Brigham and Women's Hospital, June 2019 - August 2019}
\begin{itemize}
    \item  Automated data inputs with attribute suggestion feature using React.js and AJAX
    \item  Designed data collection feature by adding REST APIs using Flask and MongoDB.
    \item  Wrote about 100 unit test cases in Mocha and Chai and increased test coverage by 5\%.
\end{itemize}